\documentclass[11pt,a4paper,final]{report}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\author{Nick Fusaro}
\title{Weapons and Armour: Brands and Implications}
\begin{document}
\begin{abstract}
\section{Weapon Brands}
A brand is a special characteristic applied to a weapon. Only one can be applied to any weapon at a time. They take the form of increasing or decreasing character or weapon attributes, adding a special on-hit or on-crit effect to the weapon, or adding an e(V)okable effect when wielded. When displayed, they follow the format "of X", applied after other weapon description modifiers. Each subsection displays the specific brand and discusses the gameplay effects of the brand briefly (if needed).
\subsection{of Protection}
Adds +5 to AC, can spawn on all weapons of all materials. Cannot spawn on armour, that would be silly.
\subsection{of Cleaving}
Causes 65 percent of damage dealt to splash to all squares adjacent to the character. Also, reduces damage dealt to swung square by 15 percent. Useful for fighters who find themselves getting caught out by many weak monsters, and in wide-open areas. Great for the tanky, who will find that the overall dps output cannot be matched by single targets, and can afford to be surrounded by many monsters that are all dying slowly.
\subsection{of Flame}
Does an additional 1d4 fire damage every successful hit. It's not particularly strong, but it does apply every time you use it. It's great for high-speed weapons, whittling down fire-weak monsters, and accurate, high-speed characters. If twoweaponing is an option, it would be good for that.
\subsection{of Venom}
Has a chance to poison a monster on-hit, of roughly 2/monster strength metric. Poison should stack to a certain degree, but we'll figure that out later. Great for fast characters, and players that prefer attrition-based gameplay. Finding one of these early on means you're basically set.
\subsection{of Frost}
Does an additional 2d4+2 frost damage 20\% of the time on-hit. Has a better expected value than flame weapons, but a greater variance. Could leave you QUITE HIGH N DRY. 
\subsection{of Speed}
Hits 1/3 faster than it usually would.
\subsection{of Power}
Critical strikes do an additional 50\% damage (additive).
\subsection{of Recursion}
Each successful critical strike is rolled a second time, with a threshold for success 20\% lower than it usually would be. Success on this adds a second crit, doubling the damage.
\subsection{of Dispersal}
Has a low on-hit chance to blink a monster. It's not really that useful except situationally.
\subsection{of Robustness}
Adjusts your HP up by 10\%.
\subsection{of Will}
Adjusts your MP up by 10\%.
\subsection{of Returning}
When thrown, has a chance to return to your quiver.
\subsection{of Impact}
Has a power 1d4+2 higher than it usually would be.
\subsection{of Haste}
Makes your entire character 1/3 faster, and significantly hungrier.
\subsection{of Evasion}
Grants +5 EV.

\section{Armor Brands}


\end{abstract}
\end{document}